---
title: "Sandbox_Cleaned"
author: "Tasaporn Visawameteekul, Law Shiang Row, Khoo Wei Lun"
format: 
  html:
    mainfont: system-ui
    fontsize: medium
    code-block-border-left: "#63B9A0"
    code-overflow: wrap
    highlight-style: solarized
    page-layout: full
    number-sections: true
    title-block-banner: true
    fig-align: center
    reader-mode: true
---

## Install and launching R packages

```{r init, results='hide', message=FALSE}
packages = c('tidyverse', 'ggstatsplot', 'psych', 'lubridate', 'ggrepel', 'plotly', "tidyr", "readr")
for(p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Read data and compress in RDS file

```{r read data, warning=FALSE, message=FALSE}
library(readr)

# Peak System Demand
T2.3 <- read_csv("data/T2-3.csv")
# Total Household Electricity Consumption by Dwelling Type
T3.4 <- read_csv("data/T3-4.csv")
# Average Monthly Household Electricity Consumption by Planning Area & Dwelling Type
T3.5 <- read_csv("data/T3-5.csv")
```

### Data table

T2.3 : Peak System Demand

```{r T2.3, echo=FALSE}
#| fig-width: 10
head(T2.3, 5)
```

T3.4 : Total Household Electricity Consumption by Dwelling Type

```{r T3.4, echo=FALSE}
#| fig-width: 10
head(T3.4, 5)
```

T3.5 : Average Monthly Household Electricity Consumption by Planning Area & Dwelling Type

```{r T3.5, echo=FALSE}
#| fig-width: 10
head(T3.5, 5)
```

## 1 Overview

### 1.1 Introduction

```{r 1.1 Introduction, warning=FALSE, message=FALSE}
library(shiny)
library(shinydashboard)

introtext = "Singapore has progressively moved towards an open electricity market since 2001 
to ensure a reliable energy supply and promote effective competition in the energy market.
However, it is challenging to analyze the energy market and make informed decisions due to 
several variable components that could influence the market. First, the price of electricity 
in the wholesale market fluctuates depending on the electricity demand and supply, which are 
dependent on consumption patterns as well as gas industry respectively. About 95% of electricity 
is generated using natural gas. Next, there are also multiple parties (Electricity Market Authority, 
Power generation companies, Electricity Retailers, etc) involved in the electricity market.
To address this challenge, we build this RShinny app to provide relevant stakeholders with means 
to analyse and understand the data with applicable analytics models. Also, we want to help the 
users explore more information about the Singapore energy market easily through visualizations."

ui = dashboardPage(
  dashboardHeader(title = 'Singapore Energy Consumption', titleWidth = 400),
  
  dashboardSidebar(width = 210,
                   sidebarMenu(
                     menuItem(" OVERVIEW", tabName = 'overview',    icon = icon("chalkboard-user"))
                   )),
  dashboardBody(
    tabItems(
      tabItem(
        tabName = "overview",
        navbarPage( "OVERVIEW",
                    
                    tabPanel("Introduction",introtext)
        )
      )
    )
  )
)
server = function(input, output, session) {}

shinyApp(ui = ui, server = server)
```

### 1.2 Geofacet

```{r 1.2 Geofacet, warning=FALSE, message=FALSE}
library(shiny)
library(shinydashboard)
library(geofacet)

town <- subset(T3.5, Description != 'Overall' & Description !='Central Region' & 
                 Description !='East Region' & Description !='North East Region' &
                 Description !='North Region' & Description !='West Region' &
                 kwh_per_acc != 's' & dwelling_type != 'Private Housing' &
                 dwelling_type != 'Public Housing' & month != 'Annual') %>%
  mutate(kwh_per_acc = as.numeric(kwh_per_acc)) %>%
  mutate(date = parse_date_time(paste0(year, "-", month,"-1"),"ymd"))


# reading the map file
mpsz <- st_read(dsn = 'master-plan-2014-subzone-boundary-web-shp',
                layer = 'MP14_SUBZONE_WEB_PL',
                crs = 3414)

# Import the area grid data.
area_grid <- read_csv("data/areagrid.csv")
singapore <- st_transform(mpsz, 4326)


ui = dashboardPage(
  dashboardHeader(title = 'Singapore Energy Consumption', titleWidth = 400),
  
  dashboardSidebar(width = 210,
                   sidebarMenu(
                     menuItem(" OVERVIEW", tabName = 'overview',    icon = icon("chalkboard-user"))
                   )),
  dashboardBody(
    tabItems(
      tabItem(
        tabName = "overview",
        navbarPage( "OVERVIEW",
                    tabPanel("Consumption by Planning Area & Dwelling Type",
                            fluidPage(
                              radioButtons("axis", label = "select axis control",
                                           choices = c("fixed y-axis" = "fixed",
                                                       "free y-axis" = "free_y"), 
                                           inline = T),
                              plotOutput("geo", height = 800)
                            ))
        )
      )
    )
  )
)

server = function(input, output, session) {
   geofacet <- town %>% 
        group_by(year, dwelling_type, Description)  %>%
        summarise(average_consumption = mean(kwh_per_acc, na.rm = TRUE))%>%
        ungroup()
  geofacet_gas_consump <- inner_join(geofacet, area_grid,
                                     by = c("Description" = "name"))
  # merge table with town name
  common_grid <- area_grid[area_grid$name %in% unique(geofacet$Description),]
  
  observeEvent(input$axis,{
    
    output$geo <- renderPlot ({
      ggplot(geofacet_gas_consump, aes(x = year, y = average_consumption)) +
        geom_line(aes(color = as.factor(dwelling_type))) +
        facet_geo(~Description, grid = common_grid, scales = input$axis) +
        labs(title = "Average Monthly Household Electricity Consumption by Planning Area & Dwelling Type") +
        theme(plot.title = element_text(size=22),
              axis.text.x = element_text(size = 10, angle = 45),
              axis.text.y = element_text(size = 10),
              strip.text = element_text(size = 10),
              legend.position = "right")
    })
  })
}

shinyApp(ui = ui, server = server)
```

### 1.3 Consumption by Dwelling type and Town

#### Plot 1 - Line plot

```{r 1.3.1 Line plot, warning=FALSE, message=FALSE}
library(tidyquant)

  sysdemand <- T2.3 %>%
    mutate(date = parse_date_time(paste0(year, "-", mth,"-1"),"ymd")) %>%
    mutate(monthyear = format(as.Date(date), "%b'%Y"))
  

    p_line <- sysdemand %>%
      mutate(text = paste(monthyear,
                          "<br>System Demand (NW):", peak_system_demand_mw)) %>%
      ggplot(aes(x = date, y = peak_system_demand_mw)) +
      geom_line() +
      geom_point() +
      labs(title = "Monthly Peak System Demand (MW)",
           x = "", y = "MW") +
      theme_tq() +
      theme(legend.position="none")
    p_line
```

#### Plot 2 - Cycle plot

```{r 1.3.2 Cycle plot, warning=FALSE, message=FALSE}
  sysdemand21 <- sysdemand %>%
    filter(year %in% c(2005:2021))
  
  #Computing year average by months
  hline.data <- sysdemand21 %>%
    group_by(mth) %>%
    summarise(avg_demand = mean(peak_system_demand_mw))
  
  #Allow user to choose monthly vs quarterly cycleplot
  p_cycleplot <-
    ggplot() +
    geom_line(data = sysdemand21,
              aes(x=year,y=peak_system_demand_mw, group=mth), colour = "black") +
    geom_hline(data = hline.data,
               aes(yintercept=avg_demand),
               linetype=6, 
               colour="red", 
               size=0.5) +
    facet_grid(~mth) +
    theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) +
    labs(title = "Cycleplot of Peak System Demand (MW), 2005 to 2021") +
    xlab("") +
    ylab("MW")
  p_cycleplot
```

#### Plot 3 - Sparkline Table

```{r 1.3.3 Sparkline table, warning=FALSE, message=FALSE}

```

#### Plot 4 - Slopegraph

```{r 1.3.4 Slope graph, warning=FALSE, message=FALSE}
#Display rate of change using Slopegraph
#Give user choice of mean, median, sum, min or max
cons_year <- chosendata %>%
  group_by(type, year) %>%
  summarise(mean_cons=round(mean(consumption),2))
p_slopegraph <- cons_year %>% 
  mutate(year = factor(year)) %>%
  filter(year %in% c(startyear,endyear)) %>%
  newggslopegraph(year, mean_cons, type)
p_slopegraph1 <- p_slopegraph + labs(title = "Average Monthly Electricity Consumption between 2 years",
                subtitle = "",
                caption = "Source: ema.gov.sg")
p_slopegraph1
```

## 2. Clustering

### 2.1 Hierachical clustering

#### Clustering data preparation

```{r 2.1.1 Method selection table, warning=FALSE, message=FALSE}
# Filter the required data for clustering

# remove month = "Annual"
# remove dewlling_type/description = Overall
# year 2018 and onwards due to missing data
# exclude "%region" in description
# Exclude Pioneer as data is incomplete

clus_data <- T3.5 %>% 
  filter(month != "Annual" & 
           year > 2017 & 
           dwelling_type != "Overall" &
           !str_detect(Description,"Region|Pioneer|Overall"))

# transform dataset
# convert kwh into numbers
clus_data$kwh_per_acc <- as.numeric(clus_data$kwh_per_acc)
# join month and year into a date
clus_data$date <- parse_date_time(paste(clus_data$year, clus_data$month), orders=c("%Y %m")) 

# drop month and year column
clus_data <- subset(clus_data, select=-c(month, year, Region)) %>%
  arrange(date)

# pivot wider
clus <- clus_data %>%
  pivot_wider(names_from=date, values_from=kwh_per_acc) 

# omit na
clus <- na.omit(clus)
clus <- clus %>% relocate(Description, .before = dwelling_type)
```

#### Plot 1 - Method selection table

```{r}
# parameter: method 
# user will need to input the distance calculated method here to determine which clustering method is optimal

library(dendextend)
clustering <- dist(normalize(clus_group1,-c(1)),  method = "euclidean")
dend_expend(clustering)[[3]]
```

#### Plot 2 - Heatmap

```{r 2.1.2 Heatmap, warning=FALSE, message=FALSE}
library(heatmaply)

# parameter 1: distance method ("euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski")
# parameter 2: hclust method ("ward.D", "ward.D2","single","complete","average", "mcquitty", "median" or "centroid")
# parameter 3: number of clusters
# parameter 4: seriate (Optimal leaf ordering, Gruvaeus and Wainer, mean, none)
# parameter 5: scale / normalize / percentize (the code for this part is different)

# Remove dwelling type
clus_group1 <- clus[,-c(2)] %>%
  group_by(Description) %>%
  summarise_each(list(sum))

# making "Description" the row name (index)
row.names(clus_group1) <- clus_group1$Description

# Making it into a matrix
clus_matrix1 <- data.matrix(clus_group1)

# plot
heatmaply(clus_matrix1[,-c(1)],
          scale = "column",
          dist_method = "euclidean",
          hclust_method = "average",
          Colv=NA,
          seriate = "none",
          k_row = 3,
          margins = c(NA,200,50,NA),
          colors = viridis(
            n= 256, alpha=1, 
            begin=0, end=1,
            option="viridis"),
          fontsize_row = 5,
          fontsize_col = 5,
          main="Hierarchical Clustering",
          ylab = "Towns",
          xlab = "Time")
```

#### Plot 3 - Number of cluster plot

```{r 2.1.3 Number of cluster plot, warning=FALSE, message=FALSE}
# once user have gotten the clustering method from above, they need to input that to find the optimal number of k
clust2 <- hclust(clustering, method = "mcquitty")
num_k <- find_k(clust2)
plot(num_k)
```

#### Plot 4 - Map

```{r 2.1.4 map, warning=FALSE, message=FALSE}

```

#### 
